---
title: "Sardine_Models_Final_04_October_2021"
author: "Abdelouahed Ben Mhamed"
date: "10/5/2021"
output: html_document
---

```{r setup, include=FALSE}
rm(list = ls())
library(knitr)
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, 
                      warning = FALSE, message = FALSE,
                      dpi = 180, fig.width = 16, fig.height = 10)
```

# Load packages

```{r}
library(janitor)
# data prep
library(dplyr)
# tidymodels
library(rsample)
library(recipes)
library(parsnip)
library(tune)
library(dials)
library(workflows)
library(yardstick)
library(readr)
library(lubridate)

```

# Read the data

## Features
```{r}
selected_features <- read.table(file = "./chinchard_features_selected_50F_PowerTrans_rsq69_rmse0.11_before_preped.txt",header = F,sep = "\n")
```

## Monthly catch
```{r}
tot_catch_per_month_all_zone <- read_csv("../data/maia_monthly_trachurus_catch_t.csv") %>% 
  clean_names() %>%
  mutate(lag_monthly_catch_t = monthly_catch_t) %>% 
  select(-monthly_catch_t)
```

## Join data
```{r}
chin_weekly <- readr::read_csv("../data/data_2015_2021_Trachurus_trachurus_weekly_10nm.csv") %>%
  clean_names() %>%
  filter(weight_t > 0) %>%
  select(all_of(selected_features$V1)) %>%
  mutate(lag_date = case_when(month(date,label = F)==1 ~ date - 31,
                              month(date,label = F)==2 & leap_year(year(date)) ~ date - 29,
                              month(date,label = F)==2 & !leap_year(year(date)) ~ date - 28,
                              month(date,label = F)==3 ~ date - 31,
                              month(date,label = F)==4 ~ date - 30,
                              month(date,label = F)==5 ~ date - 31,
                              month(date,label = F)==6 ~ date - 30,
                              month(date,label = F)==7 ~ date - 31,
                              month(date,label = F)==8 ~ date - 31,
                              month(date,label = F)==9 ~ date - 30,
                              month(date,label = F)==10 ~ date - 31,
                              month(date,label = F)==11 ~ date - 30,
                              TRUE ~ date - 31,
                              )) %>%
  mutate(year = year(date),month = month(date,label = F)) %>% 
  mutate(key =  paste(year(lag_date),month(lag_date,label = F),sep = "_")) %>%
  relocate(key,date,year,month,weight_t,lat,lon,depth) %>%
  mutate(weight_t = weight_t^0.075)

chin_weekly_with_lag <- chin_weekly %>%
  filter(date > date("2015-02-01")) %>%
  left_join(tot_catch_per_month_all_zone,by = "key") %>%
  mutate(lag_monthly_catch_t = replace(lag_monthly_catch_t, is.na(lag_monthly_catch_t),0)) %>%
  relocate(lag_monthly_catch_t, .after = weight_t) %>%
  select(-key,-year,-lag_date) %>%
  mutate(month = month(date,label = T))

rm(selected_features)
```

# Splitting the Data
```{r}
# split into training and testing datasets. Stratify by Sale price 
chin_split <- rsample::initial_split(
  chin_weekly_with_lag, 
  prop = 0.8, 
  strata = weight_t
)

chin_train <- training(chin_split)
chin_test <- testing(chin_split)

rm(chin_weekly)
rm(chin_weekly_with_lag)
rm(tot_catch_per_month_all_zone)
```

# Calculating the O2_mean for imputation

```{r}
o2_mean <- chin_train %>%
  select(names(chin_train)[grepl(pattern = "o2_mean",x = names(chin_train))])

maxP1_o2_mean_leftskew <- apply(o2_mean, 2,max,na.rm=TRUE) +1
saveRDS(maxP1_o2_mean_leftskew,file = "chin_regression_maxP1_o2_new_04Sept2021_50F.rds")
```

# Transforming O2_mean using custom methods

```{r}
# Calculating the max of each variable
f <- function(x,maxp1i){1/(maxp1i - x)}
chin_regression_maxP1_o2 <- readRDS(file = "chin_regression_maxP1_o2_new_04Sept2021_50F.rds")

chin_train <- chin_train %>%
  mutate(o2_mean9 = f(o2_mean9,chin_regression_maxP1_o2))

chin_test <- chin_test %>%
  mutate(o2_mean9 = f(o2_mean9,chin_regression_maxP1_o2))
```

# Preprocessing
```{r}
# preprocessing "recipe"
chin_rec <- recipe(weight_t ~ ., data = chin_train) %>% 
  step_mutate(datum = as.numeric(date)) %>%
  update_role(date,month,new_role = "ID") %>%
  step_knnimpute(all_predictors()) %>%
  step_inverse(depth) %>%
  step_YeoJohnson(sst3,sst21,sst42) %>%
  step_log(o2_std18) %>%
  step_sqrt(pp_mean33) %>%
  step_inverse(temp_mean3,temp_mean9,temp_mean36,temp_mean39) %>%
  step_log(temp_std9) %>%
  step_log(v_std18) %>%
  step_log(pres_std6,pres_std24,pres_std39) %>%
  step_sqrt(meteo_v_std30) %>%
  step_sqrt(hs_mean3,hs_mean15) %>%
  step_log(hs_std36) %>%
  step_ns(lat, lon, deg_free = 4)

chin_train_prep <- chin_rec %>% prep() %>% bake(new_data = NULL)
chin_test_prep <- chin_rec %>% prep() %>% bake(new_data = chin_test)
```

# Model specification
```{r}
# XGBoost model specification

chin_spec <- boost_tree(
  trees = tune(),
  tree_depth = tune(), min_n = tune(), 
  loss_reduction = tune(),sample_size = tune(), mtry = tune(),  
  learn_rate = tune(),
) %>%
  set_engine("xgboost") %>%
  set_mode("regression")
```

# Grid specification
```{r}
# grid specification
chin_grid <- grid_max_entropy(
  trees(),
  tree_depth(),min_n(),loss_reduction(),
  sample_size = sample_prop(),finalize(mtry(), chin_train),
  learn_rate(),
  size = 200
)
```

# Define the Workflow
```{r}
chin_wf <- workflow() %>%
  add_recipe(chin_rec) %>%
  add_model(chin_spec)
```

# Splitting the folds
```{r}
set.seed(234)
chin_folds <- vfold_cv(chin_train,strata = weight_t,v = 5)
```

# Tune the model

```{r}
set.seed(34567)
all_cores <- parallel::detectCores(logical = FALSE) 
library(doParallel)
cl <- makePSOCKcluster(all_cores)
registerDoParallel(cl)

chin_tune <- tune_grid(
  chin_wf,
  resamples = chin_folds,
  metrics = metric_set(rmse, rsq, mae),
  grid = chin_grid,
  control = control_grid(save_pred = TRUE,save_workflow = TRUE)
)

stopCluster(cl)
rm(cl)

save.image("all_objects_chinchard_powerTrans.RData")
```

```{r}
chin_tune %>% show_best(metric = "rsq")
```

## Select best parameters and finalize workflow
```{r}
best_params <- chin_tune %>% select_best(metric = "rmse")
xgb_workflow_finalized <- chin_wf %>% finalize_workflow(parameters = best_params)
```

## fit on all train_data and test on test_data

```{r}
chin_lastfit <- xgb_workflow_finalized %>% last_fit(chin_split)
```
## Check validation scores
```{r}
test_performance <- chin_lastfit %>% collect_metrics()
test_performance
```

# Fitting final Model
```{r}
i <- sample(x = 1:1920,size = 1)
df <- chin_train %>% bind_rows(chin_test[-i,])
final_model <- fit(xgb_workflow_finalized, df)
```

# Test if the model works
```{r}
new_data  = chin_test[i,]
predict(final_model,new_data = new_data)
```

# Save the best model
```{r}
saveRDS(object = final_model,file = "chinchard_final_best_model_workflow.rds")
```

